A Descriptive Analysis of How Psychology Journals Handle Post-Publication

### 2024-04-24 analysis/resultsAnalysis.Rmd
added article analyses

```{r}
# Load libraries
library(here)
library(lsr)
library(tidyverse)
library(tidylog)
library(binom)
library(ggplot2)
library(RColorBrewer)
library(assertthat) 
```

@@

@@

# Run a test to ensure all prominent sample and non-empirical journals removed from dj_prominent
assert_that(
  all(dj_random$sample_id == "random" | 
        dj_random$sample_id == "both" 
      & dj_random$is_empirical == TRUE),
  msg = "Found unexpected values in the sample_id or is_empirical columns"
  )

@@

Inspect how frequently post-publication critique policies were stated for prominent journals 
```{r}
prominentPPC <- dj_prominent %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

prominentPPC
```
Get the names of the prominent journals that have 'NO EXPLICIT' PPC options
```{r}
dj_prominent %>%
  filter(ppc_type == "A" & has_ppc == "NO EXPLICIT") %>%
  print()
```
Get how many prominent journals offered more than one type
```{r}
# Return all rows with ppc_name values for ppc_type B and C
dj_prominent %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name)))

@@

Inspect post-publication critique types at prominent journals
```{r}
# Counts for all types of PPC, commentary only, and letters only
promAllTypes <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  summarise(n())

@@

Inspect how frequently post-publication critique policies were stated for randomly selected journals 
```{r}
randomPPC <- dj_random %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

@@

Calculate confidence interval for explicitly stated post-publication critique policies among randomly selected journals
```{r}

@@

randomPPC_CI
```
Get how many randomly selected journals offered more than one type
```{r}
# Return all rows with ppc_name values for ppc_type B and C
dj_random %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name)))

@@

Inspect post-publication critique types at randomly selected journals
```{r}
randomTypes <- dj_random %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

@@

# What limits did journals place on post-publication critique?
Find the most restrictive limits of any journal
```{r}
# Find the lowest length limit
minLength <- d_journals %>%
 select(c(sample_id, journal_name, ppc_name, ppc_length)) %>%
 filter(!is.na(ppc_name),
         ppc_length != "NOT STATED" & ppc_length != "Qualitative") %>%
 mutate(quantLength = as.numeric(ppc_length)) %>%
 slice_min(quantLength, n = 1) %>% # Select the row with the minimum quantLength
 ungroup()

@@

## Calculations for Table 1: Prominent journals
Check length limits placed on PPC options in prominent journals
```{r}
# Summarise length statistics for All Types
prominentLengthAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_length)) %>%
 mutate(quantLength = case_when(
    ppc_length == "NOT STATED" | ppc_length == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_length) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_length == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_length == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianLength = median(quantLength, na.rm = TRUE),
    Q1 = quantile(quantLength, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantLength, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

@@

prominentLengthType
```
Check time-to-submit limits placed on PPC options in prominent journals
```{r}
# Summarise time-to-submit statistics for All Types
prominentTimeAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_time)) %>%
 mutate(quantTime = case_when(
    ppc_time == "NOT STATED" | ppc_time == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_time) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_time == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_time == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianTime = median(quantTime, na.rm = TRUE),
    Q1 = quantile(quantTime, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantTime, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

@@

prominentTimeType
```
Check reference limits placed on PPC options in prominent journals
```{r}
# Summarise reference statistics for All Types
prominentRefAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_ref)) %>%
 mutate(quantRef = case_when(
    ppc_ref == "NOT STATED" | ppc_ref == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_ref) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_ref == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_ref == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianRef = median(quantRef, na.rm = TRUE),
    Q1 = quantile(quantRef, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantRef, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

@@

prominentRefType
```
Check how many PPC options are sent for external peer review in prominent journals
```{r}
# Peer review statistics for All Types
prominentReviewAll <- dj_prominent %>%
  select(c(journal_name, ppc_name, ppc_review)) %>%
  filter(!is.na(ppc_name)) %>%
  summarise(anyReview = n() - sum(ppc_review == "NOT STATED"),
            yesReview = anyReview - sum(ppc_review == "NO"),
            yesPercent = round((yesReview / n()) * 100)   # Percentage reviewed from All Types 
  )

@@

>>>>

```{r}
# Load libraries
library(here)
library(lsr)
library(dplyr)
library(tidyverse)
library(tidylog)
library(binom)
library(ggplot2)
library(RColorBrewer)
library(assertthat) 
```

@@

*Note.* PPC = post-publication critique

@@

# Run a test to ensure all prominent sample and non-empirical journals removed from dj_random
assert_that(
  all(dj_random$sample_id == "random" | 
        dj_random$sample_id == "both" 
      & dj_random$is_empirical == TRUE),
  msg = "Found unexpected values in the sample_id or is_empirical columns"
  )

@@

Inspect how frequently PPC policies were stated for prominent journals 
```{r}
prominentPPC <- dj_prominent %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

prominentPPC
```
Get the names of the prominent journals that have 'NO EXPLICIT' PPC options
```{r}
dj_prominent %>%
  filter(ppc_type == "A" & has_ppc == "NO EXPLICIT") %>%
  print()
```
Get how many prominent journals offered more than one type of PPC
```{r}
# Return all rows with ppc_name values for ppc_type B and C
dj_prominent %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name)))

@@

Inspect PPC types at prominent journals
```{r}
# Counts for all types of PPC, commentary only, and letters only
promAllTypes <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  summarise(n())

@@

Inspect how frequently PPC policies were stated for randomly selected journals 
```{r}
randomPPC <- dj_random %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

@@

Calculate confidence interval for explicitly stated PPC policies among randomly selected journals
```{r}

@@

randomPPC_CI
```
Get how many randomly selected journals offered more than one type of PPC
```{r}
# Return all rows with ppc_name values for ppc_type B and C
dj_random %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name)))

@@

Inspect PPC types at randomly selected journals
```{r}
randomTypes <- dj_random %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

@@

# What limits did journals place on post-publication critique?

Find the most restrictive limits of any journal
```{r}
# Find the lowest length limit
minLength <- d_journals %>%
 select(c(sample_id, journal_name, ppc_name, ppc_length)) %>%
 filter(!is.na(ppc_name),
         ppc_length != "NOT STATED" & ppc_length != "Qualitative") %>%
 mutate(quantLength = as.numeric(ppc_length)) %>%
 slice_min(quantLength, n = 1) %>% # Select the row with the minimum quantLength
 ungroup()

@@

## Calculations for Table 1: Prominent journals

Check length limits placed on PPC options in prominent journals
```{r}
# Summarise length statistics for All Types
prominentLengthAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_length)) %>%
 mutate(quantLength = case_when(
    ppc_length == "NOT STATED" | ppc_length == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_length) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_length == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_length == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianLength = median(quantLength, na.rm = TRUE),
    Q1 = quantile(quantLength, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantLength, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

@@

prominentLengthType
```

Check time-to-submit limits placed on PPC options in prominent journals
```{r}
# Summarise time-to-submit statistics for All Types
prominentTimeAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_time)) %>%
 mutate(quantTime = case_when(
    ppc_time == "NOT STATED" | ppc_time == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_time) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_time == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_time == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianTime = median(quantTime, na.rm = TRUE),
    Q1 = quantile(quantTime, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantTime, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

@@

prominentTimeType
```

Check reference limits placed on PPC options in prominent journals
```{r}
# Summarise reference statistics for All Types
prominentRefAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_ref)) %>%
 mutate(quantRef = case_when(
    ppc_ref == "NOT STATED" | ppc_ref == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_ref) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_ref == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_ref == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianRef = median(quantRef, na.rm = TRUE),
    Q1 = quantile(quantRef, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantRef, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

@@

prominentRefType
```

Check how many PPC options are sent for external peer review in prominent journals
```{r}
# Peer review statistics for All Types
prominentReviewAll <- dj_prominent %>%
  select(c(journal_name, ppc_name, ppc_review)) %>%
  filter(!is.na(ppc_name)) %>%
  summarise(anyReview = n() - sum(ppc_review == "NOT STATED"),
            yesReview = anyReview - sum(ppc_review == "NO"),
            yesPercent = round((yesReview / n()) * 100)   # Percentage reviewed from All Types 
  )

@@

randomReviewAll
```
# STUDY 2: PREVALENCE

Look for links to PPC in articles published in prominent journals
```{r}
d_articles %>%
  filter(sample_id == "prominent",
         ppc_linked == TRUE) %>%
  summarise(n = n())
```

Calculate CIs for Linked PPC 
```{r}
sample_size_a <- 101

# Linked PPCs
prominent_linked <- d_articles %>% # counts number of linked PPC events
  filter(sample_id == "prominent",
         ppc_linked == TRUE) %>%
  count()

ci_linked <- binom.confint(prominent_linked$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_promLinked_LL <- round(ci_linked$lower*100, 1)
ci_promLinked_UL <- round(ci_linked$upper*100, 1)

ci_promLinked_LL
ci_promLinked_UL
```

Look for instances of PPC in articles published in prominent journals
```{r}
d_articles %>%
  filter(sample_id == "prominent",
         exclude_reason == "IS ITSELF PPC") %>%
  summarise(n = n())
```
Calculate CIs for Instance PPC in prominent journals
```{r}
# Instance PPCs
prominent_instance <- d_articles %>% # counts number of linked PPC events
  filter(sample_id == "prominent",
    exclude_reason == "IS ITSELF PPC") %>%
  count()

ci_instance <- binom.confint(prominent_instance$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_promInstance_LL <- round(ci_instance$lower*100, 1)
ci_promIinstance_UL <- round(ci_instance$upper*100, 1)

ci_promInstance_LL
ci_promIinstance_UL
```
Look for linked PPC in articles in random sample
```{r}
d_articles %>%
  filter(sample_id == "random",
         ppc_linked == TRUE) %>%
  summarise(n = n())
```

Calculate CIs for linked PPC in articles in random sample
```{r}
# Linked PPCs
random_linked <- d_articles %>% # counts number of linked PPC events
  filter(sample_id == "random",
         ppc_linked == TRUE) %>%
  count()

ci_Rlinked <- binom.confint(random_linked$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_randLinked_LL <- round(ci_Rlinked$lower*100, 1)
ci_randLinked_UL <- round(ci_Rlinked$upper*100, 1)

ci_randLinked_LL
ci_randLinked_UL
```
Look for instances of PPC in articles in random sample
```{r}
d_articles %>%
  filter(sample_id == "random",
         exclude_reason == "IS ITSELF PPC") %>%
  summarise(n = n())
```

Calculate CIs for Instance PPC in random sample
```{r}
# Instance PPCs
random_instance <- d_articles %>% # counts number of linked PPC events
  filter(sample_id == "random",
    exclude_reason == "IS ITSELF PPC") %>%
  count()

ci_Rinstance <- binom.confint(random_instance$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_randInstance_LL <- round(ci_Rinstance$lower*100, 1)
ci_randIinstance_UL <- round(ci_Rinstance$upper*100, 1)

ci_randInstance_LL
ci_randIinstance_UL
```

### 2024-04-24 manuscript/manuscript_NHB.Rmd
added article analyses

---
title: "A Descriptive Analysis of How Psychology Journals Handle Post-Publication
  Critiques"
author: Whamond, A., Vazire, S., Clarke, B., Moodie, N., Schiavone, S., Thibault,
  R. T., & Hardwicke, T. E.
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  word_document: default
abstract: "Fostering critical discussion is essential to upholding the ethos of science
  as an unbiased, collaborative, and self-correcting endeavor. Journals can support
  this by offering options for post-publication critique, such as commentaries or
  letters, allowing peers to voice substantiated concerns of empirical articles. We
  provide a descriptive, cross-sectional analysis of post-publication critique policies
  and prevalence across samples of a) 100 prominent and b) 100 randomly-selected and psychology journals.
  We found that 38% of prominent and 23% of randomly-selected psychology journals
  offered explicit critique options. Length and time-to-submit
  limits were often not stated. Critique prevalence was rare with only a single event
  identified among two samples of 101 articles. Given the widespread recognition of
  the replication crisis in psychology, allowing avenues for criticisms to be voiced
  may help increase credibility within the field. To support transparency and
  self-correction, psychology journals should offer clearly stated post-publication
  critique options and encourage their use.\n"
keywords: "post-publication critique, peer review, psychology, meta-research, journal
  policy, self-correction, scientific criticism"

bibliography: references.bib
link-citations: true
csl: nature.csl

floatsintext: false
figurelist: false
tablelist: false
footnotelist: false
linenumbers: true
mask: true
draft: true
always_allow_html: true
urlcolor: blue
---

```{r setup, include = FALSE}

library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(ggrepel) # for ggplot text labels
library(patchwork) # for multipanel plot layouts
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
library(binom) # for calculating CIs
library(citation) # for organizing citations
library(tinytex)
library(english) # for displaying code generated numbers as written 
```

@@

``` {r echo = FALSE, message = FALSE}
#ERR: Table currently inserts under results in PDF. Format is horrendous in Word.

PPCtypes %>%
  kbl(align = c(rep("l", 1), rep("c", 5), rep("l", 1)), caption = "Common post-publication critique types including median restrictions, peer review policies, and descriptions",
      col.names = c("Type", "Length", "Time-to-submit", "References", "Peer-review", "DOI", "Description")) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, align = "c")

@@

@@

# Arrange both graphs in the same figure
combined_plots <- graph3 + graph4
combined_plots
```

>>>>

---
title: "A Descriptive Analysis of How Psychology Journals Handle Post-Publication
  Critiques"
author: Whamond, A., Vazire, S., Clarke, B., Moodie, N., Schiavone, S., Thibault,
  R. T., & Hardwicke, T. E.
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  word_document: default
abstract: "Fostering critical discussion is essential to upholding the ethos of science
  as an unbiased, collaborative, and self-correcting endeavor. Journals can support
  this by offering options for post-publication critique, such as commentaries or
  letters, allowing peers to voice substantiated concerns of empirical articles. We
  provide a descriptive, cross-sectional analysis of post-publication critique policies
  and prevalence across samples of a) 100 prominent and b) 100 randomly-selected and psychology journals.
  We found that 38% of prominent and 23% of randomly-selected psychology journals
  offered explicit critique options. Length and time-to-submit
  limits were often not stated. Critique prevalence was rare with only a single event
  identified among two samples of 101 articles. Given the widespread recognition of
  the replication crisis in psychology, allowing avenues for criticisms to be voiced
  may help increase credibility within the field. To support transparency and
  self-correction, psychology journals should offer clearly stated post-publication
  critique options and encourage their use.\n"
keywords: "post-publication critique, peer review, psychology, meta-research, journal
  policy, self-correction, scientific criticism"
bibliography: references.bib
link-citations: true
csl: nature.csl
floatsintext: false
figurelist: false
tablelist: false
footnotelist: false
linenumbers: true
mask: true
draft: true
always_allow_html: true
urlcolor: blue
---

```{r setup, include = FALSE}
library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(ggrepel) # for ggplot text labels
library(patchwork) # for multipanel plot layouts
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
library(binom) # for calculating CIs
library(citation) # for organizing citations
library(tinytex)
library(english) # for displaying code generated numbers as written 
```

@@

@@

ggsave("ProminentJournalPolicies.png", plot = last_plot(), device = "png")
```

@@

# Arrange both graphs in the same figure
combined_plots <- graph3 + graph4
combined_plots

ggsave("RandomJournalPolicies.png", plot = last_plot(), device = "png")
```
